Can we distinguish modes of selective interactions using linkage disequilibrium?

### 2021-03-13 Manuscript.Rmd
add eric and matthias, some edits

In what follows, I turn to empirical measurements of LD within human
populations, focusing on mutations within protein-coding regions. I first look
at average values of LD across genes for synonymous, missense, and
loss-of-function mutations, and then turn to patterns of LD partitioned by
functional elements and distances separating mutations within genes.

Synonymous and missense mutations show similar levels of positive LD when
considering pairs of mutations within the same gene. Loss-of-function mutations
instead have negative LD, pointing to differing modes of selective interactions
for loss-of-function and missense mutations (Figure \@ref(fig:fig5)). Within
each population, large measurement noise gives 95\% confidence intervals that
overlap with zero for each mutation class, although the pattern is remarkably
consistent across African, European, and East Asian populations in the Thousand
Genomes dataset. Comparing mean LD within genes between African and Eurasian
populations, LD in Eurasian populations is somewhat larger on average, that
is, more positive for missense mutations and more negative for loss-of-function
mutations. This is in agreement with differences in expectations between
populations that have or have not gone through a bottleneck in their recent
past (Figure \@ref(fig:fig4)).

For pairs of missense mutations in different protein-coding genes, signed LD
decays to zero fairly rapidly and is lower than signed LD within genes even for
mutations that are more tightly linked (distance less that 50 kb). The
similarity in signed LD observed between missense and synonymous mutations
would suggest that interference between missense mutations is minimal. However,
below I show that interactive effects differ dramatically between
mutations that lie within annotated functional domains and mutations outside of
functional domains, with these opposing effects effectively canceling out when
averaging LD over entire genes. Due to the rarity of loss-of-function
mutations, I focus only on comparisons between synonymous and missense
mutations when looking at finer partitions of mutations within genes.

@@

To further explore the observation of positive LD between both synonymous and
missense mutations within genes, I examine the role that annotated conserved
domains in protein-coding genes play in driving this signal. Note that the low
density of loss-of-function mutations within annotated domains in humans
precludes any analysis of LD between those mutations at the level of
sub-elements within genes. Protein-coding domains are conserved elements of
genes, often associated with some known functional or structural feature of a
protein. As such, we would expect that purifying selection is stronger on
average within domains than within the same gene but outside of a conserved
domains. Indeed, the site-frequency spectrum (SFS) is skewed to lower
frequencies for both missense and loss-of-function mutations within domains
when compared to the same classes of mutations outside of domains, with much
more negative values of Tajima's D within domains (Table \@ref(tab:tajimasD)).
On the other hand, no difference is observed for synonymous mutations within or
outside domains, suggesting roughly equivalent effects of selection (either
direct or linked) on synonymous variation.

For pairs of missense mutations that both fall within the same functional
domain, we observe large positive LD, roughly an order of magnitude larger than
pairs of synonymous mutations that both fall within the same domain (Figure
\@ref(fig:fig6)). This difference between pairs of missense and synonymous
variants within domains is observed for both tightly linked ($< 10$bps) and
more distant ($>1$kb) pairs (Figure \@ref(fig:lddecaydomain)). Assuming most
synonymous mutations are neutral, we would expect signed LD between missense
mutations to be less than that between synonymous mutations under predictions
from models of either Hill-Robertson interference, synergistic epistasis, or
dominance (Figures \@ref(fig:fig1)-\@ref(fig:fig3)). This observation is
opposite to those expectations, suggesting a different prevailing interactive
effect within domains than those that are commonly tested or modeled.

The large positive signal of LD for missense mutations within the same
domain does not extend to pairs of missense mutations that are in
different domains. Pairs of missense and synonymous mutations show
nearly equal levels of LD close to zero across domains, with missense mutations
even slightly more negative than synonymous mutations (Figure \@ref(fig:fig6)).
The interactive effect driving large LD within domains is therefore likely
domain-specific. However, the average distance between pairs of mutations
within domains is much smaller than between domains, so this observation may be
primarily driven by the higher recombination rate between pairs of mutations
across distinct domains.

@@

I also measure LD between pairs of mutations that both fall outside of
annotated domains. For such mutations, I consider averages across all pairs
outside of domains, as well as subsets of pairs matched to the same distances
for pairs of mutation within and between domains (Figures \@ref(fig:fig6)B-D and
\@ref(fig:domainspops)). Averaging over all pairs of mutations within the same
gene but outside domains, both synonymous and missense mutations have LD
slightly above zero, with synonymous mutations observed to have slightly higher
signed LD than missense mutations. This pattern is primarily driven by tightly
linked variants. For pairs of mutations at distances matched to those within
the domains, synonymous mutations have much larger positive LD than
missense mutations, a signal that vanishes for pairs matched to distances
between domains. For larger distances, both synonymous and missense mutations
have LD approximately zero.

The role that tightly linked variants have in driving this signal can be
clearly seen in the decay of signed LD with distance (Figure
\@ref(fig:lddecay)). Both pairs of synonymous and missense mutations with
distances greater than a few hundred base pairs have LD that fluctuates around
zero. But for shorter distances, synonymous variants have large positive LD,
while missense mutations have much lower LD except for pairs within the
shortest distance bin ($<10$ bp). This reduction in LD is consistent with
Hill-Robertson interference between linked missense mutations instead of
epistatic effects, as interference between either additive or recessive
mutations without epistasis results in signed LD that decays rapidly by
recombination distance $\rho \approx 1$.

@@

From gene-wide averages, LD between pairs of missense mutations does not appear
to be different from pairs of synonymous variants, but LD between
loss-of-function variants is substantially more negative. Hill-Robertson
effects under additive selection are expected to be strongest for slightly to
moderately deleterious variants (with $s\sim 1/N_e$), as strongly deleterious
mutations are not expected to interfere with one another [@McVean2000-ox].
However, loss-of-function mutations do not typically fall within this regime,
and inference of the distribution of fitness effects (DFE) for new
loss-of-function variants shows that a large majority are strongly deleterious
(Supporting Information). Instead, negative synergistic epistasis between
strongly deleterious mutations does produce large negative deviations of mean
LD. Strongly deleterious *recessive* mutations can also produce this pattern,
making it difficult to distinguish these effects with certainty. This
difficulty is compounded by the large measurement noise for $\E[D]$, especially
for loss-of-function variants for which only a few hundred pairs are found
within genes in each population. If loss-of-function variants are in fact
recessive, even partially, this can impact inferences of
the DFE (Supporting Information).

@@

Positive average LD between missense and synonymous mutations has previously
been reported in humans and other species [@Sohail2017-zq; @Garcia2020-zn;
@Sandler2020-vk], and the similarity of their gene-wide LD might suggest that
interference or interactions between missense mutations are minimal, or at
least no stronger than those between synonymous mutations. However, averaging
over all observed pairs of mutations within a gene masks element-specific
interactive effects that drive LD in opposite directions. Nonsynonymous
mutations found within protein domains, identified as conserved subunits of a
gene with some structural or functional role, are more strongly selected
against on average, but also have increased signed LD over synonymous mutations
at the same distances within domains. Missense mutation outside of domains but
at the same distances as those within domains have more negative LD, both
compared to distance-matched synonymous mutations outside of domains and to
mutations within domains. Neither dominance effects, synergistic epistasis, nor
Hill-Robertson interference are expected to result in positive LD, so that some
other interactive effect must be driving this positive LD signal.

There are a number of possible interaction scenarios that can result in
positive signed LD between linked loci. One explanation is a
prevalence of pairs of compensatory mutations that are tolerated to
co-segregate at high frequencies within conserved domains [@Yeang2007-gj;
@Ivankov2014-tn]. @Callahan2011-ac and @Taverner2020-lk have proposed such a
mechanism to explain observed clusters of mutations in lineages of *Drosophila*
and other species. Another possibility is a model of antagonistic, or
diminishing returns epistasis, in which a single amino acid-changing mutation
within a domain damages the functionality of that subunit, but additional
mutations within that same domain reduces fitness by a factor less than the
first mutation. A third possibility, related to the antagonistic epistasis
case, is that selection acts on the functional domain as a unit instead of at
loci within the domain individually (such as gene-based dominance). In this
scenario, double heterozygotes have different fitnesses depending on whether
the mutations are found on same haplotype or on different haplotypes.

@@

Taken together, selection on segregating variants within protein-coding genes
is nonuniform, with both the overall strength of selection and interactions
between variants differing between annotated elements. Missense and
loss-of-function mutations within conserved domains are subject to stronger
selection, skewing the SFS to lower frequencies. However, typical approaches
for inferring the distribution of fitness effects from population genomic data
average over these differences by considering all nonsynonymous mutations
together [@Boyko2008-zk; @Kim2017-xo]. It would be straightforward to adapt
those methods to infer more detailed and accurate representation of the
heterogeneous effects of new mutations within genes by partitioning by missense
and loss-of-function classes as well as by annotated domains. Similarly, the
different modes of interaction between mutations across annotated domains mean
that our standard models and simulation approaches may be too simple and will
not capture the evolutionary trajectories and patterns of diversity that differ
at finer scales.

@@

In a single randomly mating population, neutral mutations are expected to have
average signed LD of zero, but across all populations analyzed here, LD between
synonymous mutations is positive. While selection on some subset of
synonymous variants is possible, it should be much weaker on average than
between missense mutations, and any interference between selected synonymous
variants should lead to negative LD. @Sandler2020-vk proposed that spatial
population structure may be responsible for the increase of LD for both synonymous
and missense variants. They argued that every natural population lives in some
geography that will necessarily induce some structure, and that migration, admixture,
or spatial structure could increase signed LD above zero.

However, under standard population genetic assumptions of constant mutation and
recombination rates and independent mutation events, migration and admixture
does not result in $\E[D]$ to be nonzero. For a given set of mutations with
differences in allele frequencies between source populations, $D$ will be
nonzero in an admixed population between those mutations even when $D$ is zero
in the sources [@Cavalli-Sforza1971-jv]. However, genome averages of allele
frequency differences in the source populations are zero, as they can be either
positive or negative, and resulting $\E[D]$ is still zero genome-wide.
Furthermore, @Sandler2020-vk used forward simulations with an explicit spatial
model and a multiplicative fitness function to explore this effect, and the
positive LD they observed under these scenarios extends over larger distances
than observed in the data. Instead, positive LD between synonymous mutations
decays to zero within just a few hundred base (Figures \@ref(fig:lddecay) and
\@ref(fig:lddecaydomain)).

Nonrandom mutational processes provide an alternative explanation for positive
LD between neutral variants separated by short distances, in which clusters of
mutations occur simultaneously in the same mutational event. Such
multinucleotide mutations have been shown to affect patterns of LD over short
distances, on the order of 10s to 100s of base pairs [@Harris2014-zg], and
clustered mutational events may be common in humans [@Besenbacher2016-ac].
Indeed, a simple exponential model in which the fraction of pairs of mutations
arising arising from a multinucleotide event decays with distance fits the
observed decay of $\sigma_d^1$ between synonymous mutations (Supporting
Information). We may therefore treat positive LD as the baseline expectation
for tightly linked variants primarily due to clustered mutations, so that
subsequent selective processes and interactions cause LD to deviate from that
expectation, as seen for missense mutations across different sub-elements of
genes. It may also be more appropriate to compare the negative LD observed
between linked loss-of-function variants to that positive baseline expectation
instead of zero, so that they are more recessive or have stronger synergistic
epistasis than if the neutral expectation were zero.

@@

Instead, closed, numerical approaches to compute expectations for two-locus
statistics under a wide range of selection models, as presented here, allow us
to explore this highly parameterized space of models far more efficiently, and
it opens the possibility for performing likelihood-based inference using signed
LD or other two-locus summaries of the data. For example, inferring the joint
distribution of dominance and selection is underpowered using the SFS. Because
signed LD is sensitive to the levels of dominance (Figure \@ref(fig:fig3)),
inferring the DFE with dominance may become feasible using the joint
distribution of allele frequencies and LD.

Using a single low-order summary of signed LD, such as $\E[D]$ or $\sigma_d^1$,
is likely insufficient to confidently discriminate modes of selective
interactions between linked mutations. Both dominance and epistasis can result
in strongly negative LD, although the expected decay of LD under those
scenarios can differ. Similarly, multiple interaction models can lead to
positive signed LD. Among all interaction models, the extent of LD and rate of
its decay also depends on the underlying distribution of selection coefficients
among a class of mutations, which are unknown for a given pair of mutations, so
that we must integrate over a distribution of fitness effects. This DFE,
however, will have been inferred under a simple set of assumptions, such as
additivity and interchangeability between sites within a gene, potentially
biasing any inference using this DFE to learn about other patterns of
selection.

@@

Numerical methods inhabit the space in between expensive discrete simulations
and limited analytic solutions, providing a more efficient and practical method
to compute expectations of two-locus diversity measures under a wider range of
parameters and demographic scenarios. @Ragsdale2017-gg used a finite
differencing approach to numerically solve the two-locus diffusion equation
with additive selection at either locus, although that paper focused on the
applicability of two-locus statistics to demographic inference. More recently,
@Ragsdale2019-nt extended the @Hill1968-vu system for $\E[D^2]$ to compute
arbitrary moments of the distribution of $D$ for any number of populations
connected by migration and admixture. This paper also showed that such a
moments-construction can be used to solve for the two-locus sampling
distribution within a single population, though it requires a moment-closure
approximation for nonzero recombination and selection. Below, I extend this
approach to model arbitrary diploid selection, which encompasses dominance,
epistasis, and other forms of selective interactions between two loci.

@@

This requirement of extra lineages for nonzero recombination and selection
means that the system in \@ref(eq:system) is not closed, so that we need a
moment-closure approximation to solve for $\Psi_n$. As in @Ragsdale2019-nt, a
jackknife approximation is used to estimate $\Psi_{n+l}$, for $l=1$ or $2$,
from $\Psi_n$ [@Jouganous2017-pq], so that $\hat{\Psi}_{n+l}(i, j, k) =
\mathcal{J}_{n, l}\Psi_n$, although other closure approximations are possible.
This emits a closed approximate system,
\begin{equation}
(\#eq:closed-system)
\dot{\hat\Psi}_n(i, j, k; t) = 
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}

>>>>

In what follows I examine LD from human population sequencing data, focusing on
mutations within protein-coding regions. I first look at average values of LD
across genes for pairs of synonymous, missense, and loss-of-function mutations
and then turn to patterns of LD partitioned by functional elements and
distances separating mutations within genes.

Synonymous and missense mutations show similar levels of positive LD when
considering pairs of mutations within the same gene. Loss-of-function mutations
instead have more negative LD, suggesting differing modes of selective
interactions for loss-of-function and missense mutations (Figure
\@ref(fig:fig5)). Within each population, large measurement noise gives 95\%
confidence intervals that overlap with zero across mutation class, although the
pattern is remarkably consistent across African, European, and East Asian
populations in the Thousand Genomes dataset. Comparing mean LD within genes in
African and Eurasian populations, LD in Eurasian populations is somewhat larger
on average, that is, more positive for missense mutations and more negative for
loss-of-function mutations. This is in agreement with differences in
expectations between populations that have or have not gone through a
bottleneck in their recent past (Figure \@ref(fig:fig4)). For pairs of missense
mutations in different protein-coding genes, signed LD decays to zero fairly
rapidly and is lower than signed LD within genes even for mutations that are
linked (distance less that 50 kb).

The similarity in signed LD observed between missense and synonymous mutations
would suggest that interference between missense mutations is minimal. However,
below I show that interactive effects differ dramatically between mutations
that lie within annotated functional domains and mutations outside of
functional domains, with these opposing effects effectively canceling out when
averaging LD over entire genes. Due to the rarity of loss-of-function
mutations, I focus only on comparisons between synonymous and missense
mutations when looking at finer partitions of mutations within genes.

@@

To further explore the observation of positive LD between both synonymous and
missense mutations within genes, I examine the role that annotated conserved
domains in protein-coding genes play in driving this signal. Note that the low
density of loss-of-function mutations within annotated domains in humans
precludes analysis of LD between those mutations at the level of sub-elements
within genes. Protein-coding domains are conserved elements of genes, often
associated with some known functional or structural feature of a protein. As
such, we would expect that purifying selection is stronger on average within
domains than within the same gene but outside of a conserved domains. Indeed,
the site-frequency spectrum (SFS) is skewed to lower frequencies for both
missense and loss-of-function mutations within domains when compared to the
same classes of mutations outside of domains, with much more negative values of
Tajima's D within domains (Table \@ref(tab:tajimasD)). On the other hand, no
difference is observed for synonymous mutations whether within or outside
domains, suggesting roughly equivalent effects of selection (either direct or
linked) on synonymous variation.

For pairs of missense mutations that both fall within the same functional
domain, we observe large positive LD, roughly an order of magnitude larger than
pairs of synonymous mutations that both fall within the same domain (Figure
\@ref(fig:fig6)). This difference between pairs of missense and synonymous
variants within domains is observed for both tightly linked ($< 10$bps) and
more distant ($>1$kb) pairs (Figure \@ref(fig:lddecaydomain)). Assuming most
synonymous mutations are neutral, we would expect signed LD between missense
mutations to be less than that between synonymous mutations under predictions
from models of either Hill-Robertson interference, synergistic epistasis, or
dominance (Figures \@ref(fig:fig1)-\@ref(fig:fig3)). This observation is
opposite to those expectations, suggesting a different prevailing interactive
effect between nonsynonymous mutations within domains than those that are
commonly tested or modeled.

The large positive signal of LD for missense mutations within the same
domain does not extend to pairs of missense mutations that are in
different domains. Pairs of missense and synonymous mutations show
nearly equal levels of LD close to zero across domains, with missense mutations
even slightly more negative than synonymous mutations (Figure \@ref(fig:fig6)).
The interactive effect driving large LD within domains is therefore likely
domain-specific. However, the average distance between pairs of mutations
within domains is much smaller than between domains, so this observation may be
primarily driven by the higher recombination distances between pairs of mutations
across distinct domains.

@@

I also measure LD between pairs of mutations that both fall outside of
annotated domains. For such mutations, I consider averages across all pairs of
mutations outside of domains, as well as subsets of pairs matched to the same
distances as pairs of mutation within and between domains as surveyed in the
previous section (Figures \@ref(fig:fig6)B-D and \@ref(fig:domainspops)).
Averaging over all pairs of mutations within the same gene but outside domains,
both synonymous and missense mutations have LD slightly above zero, with
synonymous mutations observed to have slightly higher signed LD than missense
mutations. This pattern is primarily driven by tightly linked variants. For
pairs of mutations at distances matched to those within the domains, synonymous
mutations have much larger positive LD than missense mutations, a signal that
vanishes for pairs matched to distances between domains. For larger distances,
both synonymous and missense mutations have LD approximately zero.

The role that tightly linked variants have in driving this signal can be
clearly seen in the decay of signed LD with distance (Figure
\@ref(fig:lddecay)). Both pairs of synonymous and missense mutations with
distances greater than a few hundred base pairs have average LD that fluctuates
around zero. But for shorter distances, synonymous variants have large positive
LD, while missense mutations have much lower LD except for pairs within the
shortest distance bin ($<10$ bp). This reduction in LD for missense mutations
compared to synonymous mutations is consistent with Hill-Robertson interference
between linked selected mutations instead of epistatic effects, as interference
between either additive or recessive mutations but without epistasis results in
signed LD that decays rapidly by recombination distances of $\rho \approx 1$.

@@

From gene-wide averages, LD between pairs of missense mutations does not appear
to be different from pairs of synonymous variants, but LD between
loss-of-function variants is substantially more negative. Hill-Robertson
effects under additive selection are expected to be strongest for slightly to
moderately deleterious variants (with $s\sim 1/N_e$), as strongly deleterious
mutations are not expected to interfere with one another [@McVean2000-ox].
However, loss-of-function mutations do not typically fall within this regime,
and inference of the distribution of fitness effects (DFE) for new
loss-of-function variants shows that a large majority are strongly deleterious
(Supporting Information). Instead, negative synergistic epistasis between
strongly deleterious mutations does produce large negative deviations of mean
LD. Strongly deleterious *recessive* mutations can also produce this pattern,
making it difficult to distinguish these effects with certainty. This
difficulty is compounded by the large measurement noise for $\E[D]$, especially
for loss-of-function variants for which only a few hundred within-gene pair in
human populations.

@@

Positive average LD between missense and synonymous mutations has previously
been reported in humans and other species [@Sohail2017-zq; @Garcia2020-zn;
@Sandler2020-vk], and the similarity of their gene-wide LD might suggest that
interference or interactions between missense mutations are minimal, or at
least no stronger than those between synonymous mutations. However, averaging
over all observed pairs of mutations within a gene masks element-specific
interactive effects that drive LD in opposite directions. Nonsynonymous
mutations found within protein domains, identified as conserved subunits of a
gene with some structural or functional role, are more strongly selected
against on average, but also have increased signed LD over synonymous mutations
at the same distances within domains. Missense mutation outside of domains but
at the same distances as those within domains have more negative LD, both
compared to distance-matched synonymous mutations outside of domains and to
mutations within domains. Neither dominance effects, synergistic epistasis, nor
Hill-Robertson interference are expected to result in positive LD, so that some
other interactive effect must be driving this positive LD signal within
conserved domains.

There are a number of possible interaction scenarios that can result in
positive signed LD between linked loci. One explanation is a prevalence of
pairs of compensatory mutations that are tolerated to co-segregate at high
frequencies within conserved domains [@Yeang2007-gj; @Ivankov2014-tn].
@Callahan2011-ac and @Taverner2020-lk have proposed such a mechanism to explain
observed clusters of mutations in lineages of *Drosophila* and other species.
Another possibility is a model of antagonistic, or diminishing returns
epistasis, in which a single amino acid-changing mutation within a domain
damages the functionality of that subunit, but additional mutations within that
same domain reduce fitness by a factor less than the first mutation. A third
possibility, related to the antagonistic epistasis case, is that selection acts
on the functional domain as a unit instead of at loci within the domain
individually (such as gene-based dominance). In this scenario, double
heterozygotes have different fitnesses depending on whether the mutations are
found on same haplotype or on different haplotypes.

@@

Taken together, selection on segregating variants within protein-coding genes
is nonuniform, with both the overall strength of selection and interactions
between variants differing between annotated elements. Missense and
loss-of-function mutations within conserved domains are subject to stronger
selection, skewing the SFS to lower frequencies. Typical approaches for
inferring the distribution of fitness effects from population genomic data
average over these differences by considering all nonsynonymous mutations
together [@Boyko2008-zk; @Kim2017-xo], but if loss-of-function variants are in
fact recessive, even partially, this can impact inferences of the DFE
(Supporting Information). It would be straightforward to adapt DFE-inference
methods to infer more detailed and accurate representation of the heterogeneous
effects of new mutations within genes by partitioning by missense and
loss-of-function classes as well as by annotated domains. Similarly, the
different modes of interaction between mutations across annotated domains mean
that our standard models and simulation approaches may be too simple and will
not capture the evolutionary trajectories and patterns of diversity that differ
at finer scales.

@@

In a single randomly mating population, neutral mutations are expected to have
average signed LD of zero, but across all populations analyzed here, LD between
synonymous mutations is positive. While selection on some subset of synonymous
variants is possible, it should be much weaker on average than between missense
mutations, and any interference between selected synonymous variants should
lead to negative LD. @Sandler2020-vk proposed that spatial population structure
may be responsible for the increase of LD for both synonymous and missense
variants. They argued that every natural population lives in some geography
that will necessarily induce some structure, and that migration, admixture, or
spatial structure could increase average signed LD above zero.

Under standard population genetic assumptions of constant mutation and
recombination rates and independent mutation events, however, migration and
admixture do not result in nonzero $\E[D]$. Rather, for a set of mutations with
*given* differences in allele frequencies between source populations, $D$ can
be nonzero in an admixed population between those particular mutations even
when $D$ is zero in the source populations [@Cavalli-Sforza1971-jv]. But this
result relies on conditioning on directional differences in allele frequencies
between populations. Expected mean allele frequency difference in the source
populations is zero ($\E[(p_1-p_2)(q_1-q_2)] = 0$), as differences are equally
likely to be positive or negative, and resulting $\E[D]$ is still zero when
taking genome-wide averages. Furthermore, @Sandler2020-vk used forward
simulations with an explicit spatial model and a multiplicative fitness
function to explore the joint effect of structure and interference, and the
positive LD they observed under these scenarios extends over much larger
distances than observed in the data. Instead, positive LD between synonymous
mutations decays to zero within just a few hundred base (Figures
\@ref(fig:lddecay) and \@ref(fig:lddecaydomain)).

Nonrandom mutational processes provide an alternative explanation for positive
LD between neutral variants separated by short distances, in which clusters of
mutations occur simultaneously in the same mutational event. Such
multinucleotide mutations have been shown to affect patterns of LD over short
distances, on the order of 10s to 100s of base pairs [@Harris2014-zg], and
clustered mutational events may be common in humans [@Besenbacher2016-ac].
Indeed, a simple exponential model in which the fraction of pairs of mutations
arising from a multinucleotide mutation event decays with distance fits the
observed decay of $\sigma_d^1$ between synonymous variants (Supporting
Information). We may therefore treat positive LD as the baseline expectation
for tightly linked variants primarily due to clustered mutations, so that
subsequent selective processes and interactions cause LD to deviate from that
expectation, as seen for missense mutations across different sub-elements of
genes. It may also be more appropriate to compare the negative LD observed
between linked loss-of-function variants to that positive baseline expectation
instead of zero, which would imply that they are more recessive or have
stronger synergistic epistasis than from inferences assuming a neutral
expectation of zero.

@@

Instead, closed, numerical approaches to compute expectations for two-locus
statistics under a wide range of selection models, as presented here or in
@Ragsdale2019-nt and @Friedlander2021-bs, allow us to explore this highly
parameterized space of models far more efficiently, and it opens the
possibility for performing likelihood-based inference using signed LD or other
two-locus summaries of the data. For example, inferring the joint distribution
of dominance and selection is underpowered using the SFS alone, but because
signed LD is sensitive to the levels of dominance (Figure \@ref(fig:fig3)),
inferring the DFE with dominance may be feasible using the joint distribution
of allele frequencies and LD.

Using a single low-order summary of signed LD, such as $\E[D]$ or $\sigma_d^1$,
is likely insufficient to confidently discriminate modes of selective
interactions between linked mutations. Both dominance and epistasis can result
in strongly negative LD, although the expected decay of LD under those
scenarios can differ. Similarly, multiple interaction models can lead to
positive signed LD. Among all interaction models, the extent of LD and rate of
its decay also depends on the underlying distribution of selection coefficients
among a class of mutations, which are unknown for a given pair of mutations, so
that we must integrate over a distribution of fitness effects. This DFE,
however, will have been inferred under a simple set of assumptions, such as
additivity and interchangeability between sites within a gene, potentially
biasing any inference using previously inferred DFEs to learn about patterns of
interactions. Again, this underlies the need to jointly infer strengths and
interactions of selected variants.

@@

Numerical methods inhabit the space in between expensive discrete simulations
and limited analytic solutions, providing a more efficient and practical method
to compute expectations of two-locus diversity measures under a wider range of
parameters and demographic scenarios. @Ragsdale2017-gg used a finite
differencing approach to numerically solve the two-locus diffusion equation
with additive selection at either locus, although that paper focused on the
applicability of two-locus statistics to demographic inference. More recently,
@Ragsdale2019-nt extended the @Hill1968-vu system for $\E[D^2]$ to compute
arbitrary moments of the distribution of $D$ for any number of populations
connected by migration and admixture. They also showed that such a
moments-construction can be used to solve for the two-locus sampling
distribution within a single population, though it requires a moment-closure
approximation for nonzero recombination and selection. Below, I extend this
approach to model arbitrary diploid selection, which encompasses dominance,
epistasis, and other forms of selective interactions between two loci. In a
concurrent study to this paper, @Friedlander2021-bs developed numerical
solutions to the same moments system, which they used to describe selected
haplotype trajectories and the distortion of neutral diversity at loci variably
linked to beneficial alleles that sweep to high frequencies under
non-equilibrium demography.

@@

This requirement of extra lineages for nonzero recombination and selection
means that the system in \@ref(eq:system) is not closed, so that we need a
moment-closure approximation to solve for $\Psi_n$. As in @Ragsdale2019-nt, a
jackknife approximation is used to estimate $\Psi_{n+l}$, for $l=1$ or $2$,
from $\Psi_n$ [@Jouganous2017-pq], so that $\hat{\Psi}_{n+l}(i, j, k) =
\mathcal{J}_{n, l}\Psi_n$, although other accurate closure approximations are
possible [@Friedlander2021-bs]. This emits a closed approximate system,
\begin{equation}
(\#eq:closed-system)
\dot{\hat\Psi}_n(i, j, k; t) = 
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}
